---
title: "R Notebook"
output: html_document
---

```{r}
library(tidyverse)
library(Seurat)
library(parallel)
library(RANN)
source("mixing_score_functions.R")
```
            

```{r}
bcr_cpg <- readRDS("data/cpg_processed.rds")
```

```{r}
DimPlot(bcr_cpg, reduction = "umap", group.by = "celltype") + coord_fixed(ratio = 1)
```



16% cells have no clone assigned.  

```{r}
sum(bcr_cpg@meta.data$clone_id == "No_contig") / ncol(bcr_cpg)
```

<!-- ```{r} -->
<!-- c = "B_60_9_2_170_2_1" -->
<!-- b = "B_126_7_4_307_2_11" -->
<!-- ``` -->



<!-- ```{r} -->
<!-- c = "B_345_7_1_167_2_1" -->
<!-- b = "B_184_5_7_167_2_1" -->
<!-- b = "B_72_4_9_204_3_1" -->
<!-- DimPlot(bcr_cpg, reduction = "UMAP", cells.highlight = list("B_345_7_1_167_2_1" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == c,]), "B_72_4_9_204_3_1" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == b,])), cols.highlight = c("blue", "red")) + coord_fixed(ratio = 1) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- DimPlot(bcr_cpg, reduction = "umap", group.by = "celltype") + coord_fixed(ratio = 1) -->
<!-- # 2 transcription factors for plasma cells -->
<!-- FeaturePlot(bcr_cpg, reduction = "umap", features = c("Xbp1", "Sdc1")) -->
<!-- ``` -->



Calculate inter-clonal mixing score. 

Cummulative sum of clones given clone size threshold. 

```{r}
bcr_cpg@meta.data %>%
  filter(clone_id != "No_contig") %>%
  count(clone_id) %>%
  group_by(n) %>%
  count(n) %>%
  # clones smaller than 3 don't make sense anyways
  filter(n >= 3) %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  mutate(cumsum = cumsum(nn)) %>%
  ggplot(aes(x=n, y=cumsum)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label=cumsum), position=position_dodge(width=0.9), vjust=-0.25) +
  theme_bw()
  
```


14s for 100 clones. 
27s for 150 clones.

The clone names must be valid R variables for the function to be fast. 

```{r}
reduction = "pca"
dims=seq(1, 20)
k = 0.3

clones <- bcr_cpg@meta.data %>%
  filter(clone_id != "No_contig") %>%
  count(clone_id) %>%
  arrange(desc(n)) %>%
  filter(n >= 8) %>%
  pull(clone_id)

# start <- Sys.time()
# set.seed(1)
# df_mc <- mixingCoefficient(object, reduction, dims, k, clones, cores=5, rep=3, clone_id = "clone_id")
# end <- Sys.time()
# end - start
```

```{r}
start <- Sys.time()
set.seed(1)
df_ms <-
  MixingScoreSeurat(
    object = bcr_cpg,
    clones = clones,
    clone_id = "clone_id",
    reduction = reduction,
    dims = dims,
    k = k,
    cores = 5
  )
end <- Sys.time()
end - start
```


```{r}
# diff_methods <- abs(df_mc[rownames(df_ms), colnames(df_ms)] - df_ms)
# which(diff_methods == max(diff_methods, na.rm = T))
```

```{r}
# hist(as.matrix(df_mc), breaks = 50)
hist(as.matrix(df_ms), breaks = 50)
```


Random background of mixing score

```{r}
bcr_cpg@meta.data$clone_id_shuffled <- bcr_cpg@meta.data$clone_id
clone_det <- bcr_cpg@meta.data$clone_id_shuffled != "No_contig"

bcr_cpg@meta.data[clone_det, "clone_id_shuffled"] <- sample(bcr_cpg@meta.data[clone_det, "clone_id_shuffled"], replace = F)

start <- Sys.time()
set.seed(1)
# df_shuffled_mc <-
#   mixingCoefficient(bcr_cpg,
#                     reduction,
#                     dims,
#                     k,
#                     clones,
#                     cores = 5,
#                     rep = 3,
#                     clone_id = "clone_id_shuffled")

df_shuffled_ms <-
  MixingScoreSeurat(
    object = bcr_cpg,
    clones = clones,
    clone_id = "clone_id_shuffled",
    reduction = reduction,
    dims = dims,
    k = k,
    cores = 5
  )

end <- Sys.time()
end - start
```


```{r}
# hist(as.matrix(df_shuffled_mc))
hist(as.matrix(df_shuffled_ms))
```

Calculate intra-clonal heterogeneity score.

Score for intra-clonal heterogeneity: mean, median and standard deviation of distance between cells in a clone.

```{r}
reduction = "pca"
dims=seq(1, 20)
object=bcr_cpg

clones <- bcr_cpg@meta.data %>%
  filter(clone_id != "No_contig") %>%
  count(clone_id) %>%
  arrange(desc(n)) %>%
  filter(n >= 8) %>%
  pull(clone_id)
```

```{r}
intra_clone_distances <- intra_clonal_distance(bcr_cpg, reduction, dims, clones)
```

Check correlation of intra-clonal heterogeneity with clone size.

```{r}
intra_clone_distances_df <- as.data.frame(t(as.data.frame(intra_clone_distances)))
colnames(intra_clone_distances_df) <- c("mean", "median", "sd")
```

```{r}
intra_clone_distances_df <-
  merge(
    intra_clone_distances_df,
    count(bcr_cpg@meta.data, clone_id, name = "clone_size"),
    by.x = "row.names",
    by.y = "clone_id"
  )
rownames(intra_clone_distances_df) <-
  intra_clone_distances_df$Row.names
intra_clone_distances_df$Row.names <- NULL
```



Find most and least homogeneous clone.

```{r}
# which(mean_intra_clone_distances == max(mean_intra_clone_distances))
# which(mean_intra_clone_distances == min(mean_intra_clone_distances))
```


```{r}
# write.csv(df, "results/mixing_coefficient_cpg_bc1_top30.csv")
# write.csv(df, "results/mixing_coefficient_cpg_top50.csv")
```


```{r}
h <- heatmap(as.matrix(df_ms), symm = T, cexRow = 0.4, cexCol = 0.4, keep.dendro=TRUE)
```

```{r}
row.clusters = as.hclust(h$Rowv)
cluster_anno <- cutree(row.clusters, k=5)  # break into k=3 clusters
```


```{r}
# df %>%
#   rownames_to_column() %>%
#   pivot_longer(cols = -rowname) %>%
#   filter(value != 1) %>%
#   arrange(desc(value)) %>%
#   head()
```


Combine intra-clonal heterogeneity with inter-clonal mixing score. 

```{r}
df_homogeneity_similarity <- df_ms %>%
  rownames_to_column(var = "clone1") %>%
  pivot_longer(cols = -clone1,
               names_to = "clone2",
               values_to = "mixing_coefficient") %>%
  filter(mixing_coefficient != 1)

df_homogeneity_similarity <-
  merge(
    df_homogeneity_similarity,
    as.data.frame(intra_clone_distances_df),
    by.x = "clone1",
    by.y = "row.names",
    all.x = T
  )
df_homogeneity_similarity <-
  rename(
    df_homogeneity_similarity,
    mean_clone1 = "mean",
    median_clone1 = "median",
    sd_clone1 = "sd",
    
  )

df_homogeneity_similarity <-
  merge(
    df_homogeneity_similarity,
    as.data.frame(intra_clone_distances_df),
    by.x = "clone2",
    by.y = "row.names",
    all.x = T
  )
df_homogeneity_similarity <-
  rename(
    df_homogeneity_similarity,
    mean_clone2 = "mean",
    median_clone2 = "median",
    sd_clone2 = "sd",
    size_clone1 = "clone_size.x",
    size_clone2 = "clone_size.y"
  )
```

Get mean PC space for top50 clones

```{r}
cells_df <-
  object@meta.data %>% filter(clone_id %in% clones) %>% select(clone_id)
cells_df$cell_id <- rownames(cells_df)

embeddings <- Embeddings(object = object[[reduction]])[, dims]

clone_pc_means <- as.data.frame(embeddings) %>%
  rownames_to_column(var = "cell_id") %>%
  merge(cells_df, by="cell_id") %>%
  group_by(clone_id) %>%
  summarise(across(where(is.numeric), mean)) %>%
  as.data.frame()

rownames(clone_pc_means) <- clone_pc_means$clone_id
clone_pc_means$clone_id <- NULL

clone_pc_means_dist <- as.data.frame(as.matrix(dist(clone_pc_means)))
```

```{r}
clone_pc_means_dist_df <- clone_pc_means_dist %>%
  rownames_to_column(var = "clone1") %>%
  pivot_longer(cols = -clone1, names_to = "clone2", values_to = "pca_distance") %>%
  filter(pca_distance != 1)

df_homogeneity_similarity <- merge(df_homogeneity_similarity, clone_pc_means_dist_df, by=c("clone1", "clone2"))
head(df_homogeneity_similarity)
```

```{r}
df_homogeneity_similarity <- df_homogeneity_similarity %>%
  mutate(median_diff = log(median_clone1 / median_clone2))

df_homogeneity_similarity <- df_homogeneity_similarity %>%
  mutate(abs_median_diff = abs(median_diff))

df_homogeneity_similarity <- df_homogeneity_similarity %>%
  mutate(mean_median = (median_clone1 + median_clone2) / 2)

df_homogeneity_similarity <- df_homogeneity_similarity %>%
  mutate(size_diff = log(size_clone1 / size_clone2))

df_homogeneity_similarity <- df_homogeneity_similarity %>%
  mutate(abs_size_diff = abs(size_diff))

df_homogeneity_similarity <- df_homogeneity_similarity %>%
  mutate(mean_size = (size_clone1 + size_clone2) / 2)
```


## kmeans on PC centroids

```{r}
cells_df <- object@meta.data %>% filter(clone_id %in% clones) %>% select(clone_id)
cells_df$cell_id <- rownames(cells_df)

embeddings <- Embeddings(object = object[[reduction]])[, dims]

clone_pc_means <- as.data.frame(embeddings) %>%
  rownames_to_column(var = "cell_id") %>%
  merge(cells_df, by="cell_id") %>%
  group_by(clone_id) %>%
  summarise(across(where(is.numeric), mean)) %>%
  as.data.frame()

rownames(clone_pc_means) <- clone_pc_means$clone_id
clone_pc_means$clone_id <- NULL
```


```{r}
clone_pc_means_kmeans <- t(clone_pc_means)
colnames(clone_pc_means_kmeans) <- rownames(clone_pc_means)

set.seed(21)
clone_pc_means_kmeans <- kmeans(x = clone_pc_means, centers = 5)
```

Add clusters to seurat metadata.

```{r}
pc_means_kmeans <- bcr_cpg@meta.data %>%
  merge(clone_pc_means_kmeans$cluster, by.x = "clone_id", by.y = "row.names", all.x=T) %>%
  rename("pc_means_kmeans" = y) %>%
  select(cell_id, pc_means_kmeans)

rownames(pc_means_kmeans) <- pc_means_kmeans$cell_id

# bcr_cpg$pc_means_kmeans <- NULL
bcr_cpg <- AddMetaData(bcr_cpg, pc_means_kmeans)

bcr_cpg@meta.data$mixing_score_kmeans <- NULL
mixing_score_kmeans <- bcr_cpg@meta.data %>%
  merge(cluster_anno, by.x = "clone_id", by.y = "row.names", all.x=T) %>%
  rename("mixing_score_kmeans" = y) %>%
  select(cell_id, mixing_score_kmeans)

rownames(mixing_score_kmeans) <- mixing_score_kmeans$cell_id

bcr_cpg <- AddMetaData(bcr_cpg, mixing_score_kmeans)
```

How to clusters from mean distance and mixing score overlap?

```{r}
bcr_cpg@meta.data %>%
  select(clone_id, pc_means_kmeans, mixing_score_kmeans) %>%
  unique() %>%
  select(pc_means_kmeans, mixing_score_kmeans) %>%
  table()
```


===========

```{r}
GGally::ggpairs(
  intra_clone_distances_df,
  aes(alpha = 0.2),
  lower = list(
    combo = GGally::wrap("dot", alpha = 0.4, size = 0.5),
    continuous = GGally::wrap("smooth", alpha = 0.3, size = 0.5)
  ),
  progress = F
)
```

```{r fig.height=8}
GGally::ggpairs(
  df_homogeneity_similarity[c(3, 12, 14, 15, 17, 18)],
  progress = F,
  lower = list(
    # continuous = GGally::wrap("points", alpha = 0.3, size = 0.1),
    # combo = GGally::wrap("dot", alpha = 0.4, size = 0.2),
    continuous = GGally::wrap("smooth", alpha = 0.05, size = 0.1)
  ),
  upper = list(continuous = GGally::wrap("cor", method = "pearson"))
) +
  theme_bw()
# ggsave("results/pairs_plot_top150.pdf")
```


```{r fig.height=6}
GGally::ggpairs(
  df_homogeneity_similarity[c(3, 12, 14, 17)],
  progress = F,
  lower = list(
    # continuous = GGally::wrap("points", alpha = 0.1, size = 0.2),
    # combo = GGally::wrap("dot", alpha = 1, size = 0.2),
    continuous = GGally::wrap("smooth", alpha = 0.05, size = 0.1)
  ),
  upper = list(continuous = GGally::wrap("cor", method = "pearson"))
) +
  theme_bw()
# ggsave("results/pairs_plot_top150.pdf")
```


```{r}
df_homogeneity_similarity %>%
  ggplot(aes(x=mixing_coefficient, y=pca_distance, color=abs_median_diff)) +
  geom_point()
```


Clones with small distance of means can have low mixing score if e.g. large difference in intra-clonal homogeneity

Negative correlation of intra-clonal heterogeneity and and mixing score for clones with small PCA distance. 

```{r fig.height=2.5}
df_homogeneity_similarity %>%
  filter(pca_distance < 10) %>%
  ggplot(aes(x=mixing_coefficient, y=abs_median_diff)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_bw()
```



```{r}
ggplot(df_homogeneity_similarity, aes(mixing_coefficient, mean_intra_clone_distances_clone1 - mean_intra_clone_distances_clone2, color=size_clone1)) +
  geom_point() +
  theme_bw()

ggplot(df_homogeneity_similarity, aes(mixing_coefficient, size_clone1)) +
  geom_point() +
  theme_bw()

ggplot(df_homogeneity_similarity, aes(mixing_coefficient, mean_intra_clone_distances_clone1)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method = "lm")

# ggsave("results/top50_cpg.pdf")
```

```{r}
df_homogeneity_similarity %>%
  filter(mixing_coefficient > 0.9, mean_intra_clone_distances_clone1 - mean_intra_clone_distances_clone2 > 5)

df_homogeneity_similarity %>%
  filter(mixing_coefficient > 0.9 & mixing_coefficient < 0.92, mean_intra_clone_distances_clone1 - mean_intra_clone_distances_clone2 < 2)

```



```{r}
df_homogeneity_similarity %>%
  arrange(desc(mean_intra_clone_distances))
```


```{r}
mean(as.matrix(df["B_131_5_22_52_2_3",]))
mean(as.matrix(df["B_278_8_8_261_1_1",]))
```


```{r}
c <- "B_278_8_8_261_1_1"
c <- "B_184_5_7_167_2_1"
c <- "B_2_5_22_312_1_3"
b <- "B_131_5_22_52_2_3"
b <- "B_250_10_1_375_2_3"
b <- "B_26_3_3_21_2_1"

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = list(
    "B_2_5_22_312_1_3" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == c, ]),
    "B_26_3_3_21_2_1" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == b, ])
  ),
  cols.highlight = c("blue", "red")
) + coord_fixed(ratio = 1)
```

```{r}
c <- "B_143_6_6_278_2_1"
b <- "B_2_5_22_312_1_3"

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = list(
    "B_143_6_6_278_2_1" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == c, ]),
    "B_2_5_22_312_1_3" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == b, ])
  ),
  cols.highlight = c("blue", "red")
) + coord_fixed(ratio = 1)
```


```{r}
distant_clones <- df_homogeneity_similarity %>%
  arrange(mixing_coefficient) %>%
  filter(mean_median < 10) %>%
  head(1) %>%
  tail(1) %>%
  select(clone1, clone2) %>%
  unlist()


distant_clones_highlight <- lapply(distant_clones, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(distant_clones_highlight) <- distant_clones

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = distant_clones_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```

```{r}
close_clones <- df_homogeneity_similarity %>%
  arrange(desc(mixing_coefficient)) %>%
  filter(mean_median < 10) %>%
  head(1) %>%
  tail(1) %>%
  select(clone1, clone2) %>%
  unlist()


close_clones_highlight <- lapply(close_clones, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(close_clones_highlight) <- close_clones

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = close_clones_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```



```{r}
df_homogeneity_similarity
hetereogeneous_clones <- df_homogeneity_similarity %>%
  arrange(desc(mean_intra_clone_distances_clone1)) %>%
  select(clone1) %>%
  unique() %>%
  head(5) %>%
  pull(clone1)

hetereogeneous_clones_highlight <- lapply(hetereogeneous_clones, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(hetereogeneous_clones_highlight) <- hetereogeneous_clones

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = hetereogeneous_clones_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```

```{r}
homogeneous_clones <- df_homogeneity_similarity %>%
  arrange(desc(mean_intra_clone_distances_clone1)) %>%
  select(clone1) %>%
  unique() %>%
  tail(5) %>%
  pull(clone1)

homogeneous_clones_highlight <- lapply(homogeneous_clones, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(homogeneous_clones_highlight) <- homogeneous_clones

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = homogeneous_clones_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```


What is different between clones that overlap very well? 

Correlation mixing coefficient and clone size and homogeneity


homogeneity between time points
Average distance between cells within a clone


For each clone, calculate the background by splitting the clone into 2?
Background: for each clone, compare similarity score with a random selection of cells


What happens if I go not symmetric? Only cells from c, not b?


Distribution of cell types across clones

```{r fig.width=10, fig.height=5}
bcr_cpg@meta.data %>%
  filter(clone_id %in% clones) %>%
  merge(cluster_anno, by.x = "clone_id", by.y = "row.names") %>%
  group_by(clone_id, y) %>%
  count(celltype) %>%
  ggplot(aes(fill=celltype, y=n, x=clone_id)) + 
    geom_bar(position="fill", stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  facet_grid(~y, scales = "free")
```




Check where clusters are on UMAP

```{r}
require(scales)

# "#FF0000" "#80FF00" "#00FFFF" "#8000FF"
DimPlot(
  subset(bcr_cpg, subset = pc_means_kmeans %in% c("1", "2", "3", "4", "5")),
  reduction = "umap",
  dims = c(1, 2),
  group.by = "pc_means_kmeans",
  na.value = "00FFFFFF",
  cols = hue_pal()(5),
  shuffle = T
) + coord_equal()
DimPlot(
  subset(bcr_cpg, subset = mixing_score_kmeans %in% c("1", "2", "3", "4", "5")),
  reduction = "umap",
  dims = c(1, 2),
  group.by = "mixing_score_kmeans",
  na.value = "00FFFFFF",
  cols = hue_pal()(5),
  shuffle = T
) + coord_equal()
DimPlot(
  bcr_cpg,
  reduction = "umap",
  dims = c(1, 2),
  group.by = "celltype",
  shuffle = T
) + coord_equal()
```

<!-- DEG clusters by kmeans or mixing score -->


<!-- ```{r} -->
<!-- bcr_cpg@meta.data -->

<!-- Idents(bcr_cpg) <- "mixing_score_kmeans" -->
<!-- mixing_score_kmeans_markers <- FindAllMarkers(bcr_cpg, assay = "originalexp", slot = "counts") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- Idents(bcr_cpg) <- "pc_means_kmeans" -->
<!-- pc_means_kmeans_markers <- FindAllMarkers(bcr_cpg, assay = "originalexp", slot = "counts") -->
<!-- ``` -->


<!-- ```{r} -->
<!-- mixing_score_kmeans_markers %>% -->
<!--   filter(p_val_adj < 0.05, pct.1 > 0.05) %>% -->
<!--   group_by(cluster) %>% -->
<!--   slice_max(order_by = avg_log2FC, n = 10) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- pc_means_kmeans_markers %>% -->
<!--   filter(p_val_adj < 0.05, pct.1 > 0.05) %>% -->
<!--   group_by(cluster) %>% -->
<!--   slice_max(order_by = avg_log2FC, n = 10) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- mixing_score_kmeans_markers_sign <- mixing_score_kmeans_markers %>% -->
<!--   filter(p_val_adj < 0.05) -->

<!-- pc_means_kmeans_markers_sign <- pc_means_kmeans_markers %>% -->
<!--   filter(p_val_adj < 0.05) -->

<!-- mixing_score_kmeans_markers_sign[!mixing_score_kmeans_markers_sign$gene %in% pc_means_kmeans_markers_sign$gene,] %>%  -->
<!--   group_by(cluster) %>% -->
<!--   slice_max(order_by = avg_log2FC, n = 10) -->
<!-- ``` -->


#### DEG pseudo bulk

Aggregate counts by clone

```{r}
seurat_object_aggr <-
  AggregateClones(seurat_object = bcr_cpg)

clone_cluster_markers_mixing_score <- CloneGroupDeg(seurat_object_aggr, cluster_anno)
clone_cluster_markers_pc_means <- CloneGroupDeg(seurat_object_aggr, clone_pc_means_kmeans$cluster)
```



DEG with DESeq2 PC centroids

```{r}
clone_cluster_markers_pc_means %>%
  filter(p_val_adj < 0.05) %>%
  slice_max(order_by = avg_log2FC, n = 10, by=cluster)
```
```{r}
library(EnhancedVolcano)
```


```{r fig.height=10}
sapply(seq(1, 4), function(x) {
  tmp <- filter(clone_cluster_markers_pc_means, cluster == as.character(x))
  par(new=T)
  plot(EnhancedVolcano(tmp,
                  lab = tmp$gene,
                  x = 'avg_log2FC',
                  y = 'p_val_adj'))
})
```

```{r}
clone_cluster_markers_pc_means %>% filter(gene == "Vim")
```


DEG with DESeq2 Mixing score
```{r}
clone_cluster_markers_mixing_score %>%
  filter(p_val_adj < 0.05) %>%
  slice_max(order_by = avg_log2FC, n = 10, by=cluster)
```

```{r fig.height=10}
sapply(seq(1, 4), function(x) {
  tmp <- filter(clone_cluster_markers_mixing_score, cluster == as.character(x))
  par(new=T)
  plot(EnhancedVolcano(tmp,
                  lab = tmp$gene,
                  x = 'avg_log2FC',
                  y = 'p_val_adj'))
})
```

```{r}
unique(pull(filter(clone_cluster_markers_mixing_score, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene)[pull(filter(clone_cluster_markers_mixing_score, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene) %in% pull(filter(clone_cluster_markers_pc_means, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene)])
cat("\n")
genes_only_ms <- unique(pull(filter(clone_cluster_markers_mixing_score, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene)[!pull(filter(clone_cluster_markers_mixing_score, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene) %in% pull(filter(clone_cluster_markers_pc_means, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene)])
genes_only_ms
cat("\n")
genes_only_centoids <- unique(pull(filter(clone_cluster_markers_pc_means, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene)[!pull(filter(clone_cluster_markers_pc_means, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene) %in% pull(filter(clone_cluster_markers_mixing_score, p_val_adj < 0.05, abs(avg_log2FC) > 1.5), gene)])
genes_only_centoids
```


```{r}
clone_cluster_markers_mixing_score %>% filter(gene %in% genes_only_centoids) %>% mutate(p_val_adj = round(p_val_adj, 3)) %>% arrange(gene)
clone_cluster_markers_pc_means %>% filter(gene %in% genes_only_ms) %>% mutate(p_val_adj = round(p_val_adj, 3)) %>% arrange(gene)
```


<!-- ```{r} -->
<!-- clone_pc_means_dist <- as.data.frame(as.matrix(dist(clone_pc_means))) -->
<!-- colnames(clone_pc_means_dist) <- clone_pc_means$clone_id -->
<!-- rownames(clone_pc_means_dist) <- clone_pc_means$clone_id -->
<!-- ``` -->

<!-- ```{r} -->
<!-- clone_pc_means_dist_df <- clone_pc_means_dist %>% -->
<!--   rownames_to_column(var = "clone1") %>% -->
<!--   pivot_longer(cols = -clone1, names_to = "clone2", values_to = "pca_distance") %>% -->
<!--   filter(pca_distance != 1) -->

<!-- df_homogeneity_similarity <- merge(df_homogeneity_similarity, clone_pc_means_dist_df, by=c("clone1", "clone2")) -->
<!-- ``` -->

```{r fig.height=4, fig.width=5}
df_homogeneity_similarity %>%
  ggplot(aes(mixing_coefficient, pca_distance, colour=abs(mean_intra_clone_distances_clone1 - mean_intra_clone_distances_clone2))) +
  theme(legend.title = element_blank())+#"abs(difference median distance")) +
  geom_point()
```


```{r fig.width=10, fig.height=5}
bcr_cpg@meta.data %>%
  filter(clone_id %in% clones) %>%
  merge(clone_pc_means_kmeans$cluster, by.x = "clone_id", by.y = "row.names") %>%
  group_by(clone_id, y) %>%
  count(celltype) %>%
  ggplot(aes(fill=celltype, y=n, x=clone_id)) + 
    geom_bar(position="fill", stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  facet_grid(~y, scales = "free")
```


```{r}
table(merge(
  as.data.frame(clone_pc_means_kmeans$cluster),
  as.data.frame(cluster_anno),
  by = "row.names"
)[c("clone_pc_means_kmeans$cluster", "cluster_anno")]) %>%
  heatmap(scale = "none")
```



### WHat is similarity score > 1?

```{r}
dim(df_homogeneity_similarity)
df_homogeneity_similarity %>%
  arrange(desc(mixing_coefficient)) %>%
  filter(mixing_coefficient > 1)
```

```{r}
clones_larger_1 <- df_homogeneity_similarity %>%
  arrange(desc(mixing_coefficient)) %>%
  head(1) %>%
  tail(1) %>%
  select(clone1, clone2) %>%
  unlist()


clones_larger_1_highlight <- lapply(clones_larger_1, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(clones_larger_1_highlight) <- clones_larger_1

DimPlot(
  bcr_cpg,
  reduction = "UMAP",
  cells.highlight = clones_larger_1_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```


