---
title: "R Notebook"
output: html_document
---

```{r}
library(tidyverse)
library(Seurat)
library(parallel)
library(RANN)
```


```{r}
getwd()
bcr <- readRDS("./data/zenodo_download/seurat_bcr.rds")

bcr
```


Plot number of barcodes per cell

Plot size of clones

clone_id "No_contig"

```{r}
table(bcr@meta.data$orig.ident)

table(bcr@meta.data$filter_rna)
```


```{r fig.width=7, fig.height=4}
bcr@meta.data %>%
  filter(filter_rna == "False") %>%
  group_by(orig.ident) %>%
  filter(clone_id != "No_contig") %>%
  count(clone_id) %>%
  ggplot(aes(x=n, fill=orig.ident)) +
  geom_bar() +
  facet_wrap(~orig.ident) +
  theme_bw()
```

```{r}
head(bcr@meta.data)
```


```{r}
bcr@meta.data %>%
  # filter(filter_rna == "False") %>%
  filter(clone_id != "No_contig") %>%
  mutate(n_barcodes = str_count(clone_id, pattern = "\\|") + 1) %>%
  ggplot(aes(n_barcodes, total_counts)) +
  geom_point(alpha=0.02)
```

```{r}
bcr@meta.data %>%
  mutate(n_barcodes = ifelse(clone_id == "No_contig", 0, str_count(clone_id, pattern = "\\|") + 1)) %>%
  group_by(orig.ident) %>%
  count(n_barcodes) %>%
  ggplot(aes(x=n_barcodes, y=n)) +
  geom_bar(stat="identity") +
  facet_wrap(~orig.ident)
```


# Mixing coefficient


Pseudo code

Create 50 PC space

For each clone c

    For each clone b that is not clone c
    
      For each repetition r
      
        subsample smaller clone
        
        for each cell in clone c
          select k nearest neighbours n
          mixing = sum(clone_id[n] == c) / k
        
            

```{r}
bcr_cpg <- readRDS("data/cpg_processed.rds")
```

```{r}
DimPlot(bcr_cpg, reduction = "UMAP", group.by = "originalexp_snn_res.0.2") + coord_fixed(ratio = 1)
DimPlot(bcr_cpg, reduction = "UMAP", group.by = "celltype") + coord_fixed(ratio = 1)
```


16% cells have no clone assigned.  

```{r}
sum(bcr_cpg@meta.data$clone_id == "No_contig") / ncol(bcr_cpg)
```

```{r}
bcr_cpg@meta.data %>%
  count(clone_id) %>%
  arrange(desc(n))
```

```{r}
c = "B_60_9_2_170_2_1"
b = "B_126_7_4_307_2_11"
```



```{r}
c = "B_345_7_1_167_2_1"
b = "B_184_5_7_167_2_1"
b = "B_72_4_9_204_3_1"
DimPlot(bcr_cpg, reduction = "umap", cells.highlight = list("B_345_7_1_167_2_1" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == c,]), "B_72_4_9_204_3_1" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == b,])), cols.highlight = c("blue", "red")) + coord_fixed(ratio = 1)
```


```{r}
DimPlot(bcr_cpg, reduction = "umap", group.by = "celltype") + coord_fixed(ratio = 1)
# 2 transcription factors for plasma cells
FeaturePlot(bcr_cpg, reduction = "umap", features = c("Xbp1", "Sdc1"))
```


Optimizing
- calculate distances in neighbourhood graph for all cells that have a barcode, 20 nearest neighbours from pre-calculated neighbourhood graph


```{r}
mixingCoefficient <- function(object, 
                              reduction,
                              dims,
                              k,
                              clones,
                              cores = 1,
                              max_cells = 50,
                              rep = 1) {
    clones <- as.character(clones)
    
    # get embedding with specified dimensions from Seurat object
    embeddings <- Embeddings(object = object[[reduction]])[, dims]
    
    # dataframe of cell IDs and clone IDs
    cells_df <- object@meta.data %>%
      rownames_to_column(var = "cell_id") %>%
      filter(clone_id %in% clones) %>%
      select(clone_id, cell_id)
    
    # subset embedding to cells from selected clones
    embeddings <- embeddings[cells_df$cell_id, ]
    
    # calculate distance matrix between all cells of all selected clones
    dist_matrix <- dist(embeddings)
    dist_matrix <- as.matrix(dist_matrix)
    
    # create clone to cell_ids dict
    clone_to_cell_dics <- split(cells_df$cell_id, cells_df$clone_id)
    
    # prepare empty results dataframe
    df_scores <-
      data.frame(matrix(nrow = length(clones), ncol = length(clones)))
    
    clone_index_pairs <-
      as.list(as.data.frame(t(
        permutations(
          n = length(clones),
          r = 2,
          repeats.allowed = F
        )
      )))
    
    clone_index_pairs <-
      clone_index_pairs[sapply(clone_index_pairs, function(x)
        x[1] < x[2])]
    
    # cells_vector
    cells_vector <- cells_df$clone_id
    names(cells_vector) <- cells_df$cell_id
    
    # iterate over all clone pairs
    mixing_scores <-
      mclapply(clone_index_pairs, mc.cores = cores, function(clone_pair) {
        c = clones[clone_pair[1]]
        b = clones[clone_pair[2]]
        
        cells1 <- as.vector(unlist(clone_to_cell_dics[c]))
        cells2 <- as.vector(unlist(clone_to_cell_dics[b]))
        
        # iterate over sampling repetitions
        mixing_scores <- sapply(rep, function(r) {
          # sample to smaller clone size
          if (length(cells1) < length(cells2)) {
            cells2 <- sample(cells2, size = length(cells1), replace = F)
          } else if (length(cells1) > length(cells2)) {
            cells1 <- sample(cells1, size = length(cells2), replace = F)
          }
          
          cell_list <- c(cells1, cells2)
          
          dist_matrix_clones <-
            dist_matrix[cell_list, cell_list]
          
          # k is half the number of cells of one clone
          k_clone_pair <- floor(length(cell_list) * 0.5 * k)
          col_dist <- colnames(dist_matrix_clones)
          # calculate average identity
          s <- sapply(cell_list, function(cell) {
            # get nearest neighbours (2-21 to not take same cell)
            kneigh <-
              col_dist[order(dist_matrix_clones[cell, ])[seq(2, k_clone_pair + 1)]]
            sum(cells_vector[kneigh] == cells_vector[cell])
          })
          
          # calculate mixing score as in
          mixing_score <-
            1 - (sum(s) / length(cell_list) - k_clone_pair / 2) / (k_clone_pair - k_clone_pair / 2)
          return(mixing_score)
        })
        
        return(mean(mixing_scores))
      })
    
    for (i in seq(1, length(clone_index_pairs))) {
      df_scores[clone_index_pairs[[i]][1], clone_index_pairs[[i]][2]] <-
        mixing_scores[[i]]
      df_scores[clone_index_pairs[[i]][2], clone_index_pairs[[i]][1]] <-
        mixing_scores[[i]]
    }
    
    rownames(df_scores) <- clones
    colnames(df_scores) <- clones
    return(df_scores)
  }
```

Score for homogeneity: average distance (including cell itself)

```{r}
intra_clonal_distance <- function(object, reduction, dims, clones) {
  mean_intra_clone_distances <- c()
  
  embeddings <- Embeddings(object = object[[reduction]])[, dims]
  
  for (c_i in seq(length(clones))) {
    c = clones[c_i]
    
    cells <-
      object@meta.data %>% filter(clone_id %in% c) %>% rownames_to_column("cell") %>% pull(cell)
    
    # get neighbourhood graph for all cells
    nn <- nn2(data = embeddings[cells,],
              k = length(cells),
              eps = eps)
    
    
    mean_intra_clone_distances <- c(mean_intra_clone_distances, median(nn$nn.dists))
  }
  
  names(mean_intra_clone_distances) <- clones
  return(mean_intra_clone_distances)
}
```


```{r}
reduction = "pca"
dims=seq(1, 20)
object=bcr_cpg
k = 0.5

clones <- bcr_cpg@meta.data %>%
  filter(clone_id != "No_contig") %>%
  count(clone_id) %>%
  arrange(desc(n)) %>%
  head(50) %>%
  pull(clone_id)

start <- Sys.time()
set.seed(1)
df <- mixingCoefficient(object, reduction, dims, k, clones, cores=10, rep=3)
end <- Sys.time()
end - start
```

```{r}
mean_intra_clone_distances <- intra_clonal_distance(object, reduction, dims, clones)
```


```{r}
hist(as.matrix(df))
max(as.matrix(df), na.rm = T)
```



```{r}
hist(as.matrix(mean_intra_clone_distances))
```

```{r}
write.csv(df, "results/mixing_coefficient_cpg_bc1_top30.csv")
write.csv(df, "results/mixing_coefficient_cpg_top50.csv")
```


```{r}
h <- heatmap(as.matrix(df), symm = T, cexRow = 0.4, cexCol = 0.4, keep.dendro=TRUE)
```
```{r}
row.clusters = as.hclust(h$Rowv)
cluster_anno <- cutree(row.clusters, k=4)  # break into k=3 clusters
```


```{r}
df %>%
  rownames_to_column() %>%
  pivot_longer(cols = -rowname) %>%
  filter(value != 1) %>%
  arrange(desc(value))
```


Distribution of similarity scores. 

```{r}
hist(as.matrix(df))
```






Find most and least homogeneous clone.

```{r}
which(mean_intra_clone_distances == max(mean_intra_clone_distances))
which(mean_intra_clone_distances == min(mean_intra_clone_distances))
```

```{r}
count(bcr_cpg_contig@meta.data, clone_id)
```


```{r}
df_homogeneity_similarity <- df %>%
  rownames_to_column(var = "clone1") %>%
  pivot_longer(cols = -clone1, names_to = "clone2", values_to = "mixing_coefficient") %>%
  filter(mixing_coefficient != 1)

df_homogeneity_similarity <- merge(df_homogeneity_similarity, as.data.frame(mean_intra_clone_distances), by.x = "clone1", by.y="row.names")
df_homogeneity_similarity <- rename(df_homogeneity_similarity, mean_intra_clone_distances_clone1 = "mean_intra_clone_distances")

df_homogeneity_similarity <- merge(df_homogeneity_similarity, as.data.frame(mean_intra_clone_distances), by.x = "clone2", by.y="row.names")
df_homogeneity_similarity <- rename(df_homogeneity_similarity, mean_intra_clone_distances_clone2 = "mean_intra_clone_distances")

df_homogeneity_similarity <- merge(df_homogeneity_similarity, count(bcr_cpg@meta.data, clone_id, name = "size_clone1"), by.x = "clone1", by.y="clone_id")
```

```{r}
ggplot(df_homogeneity_similarity, aes(mixing_coefficient, mean_intra_clone_distances_clone1 - mean_intra_clone_distances_clone2, color=size_clone1)) +
  geom_point() +
  theme_bw()

ggplot(df_homogeneity_similarity, aes(mixing_coefficient, size_clone1)) +
  geom_point() +
  theme_bw()

ggplot(df_homogeneity_similarity, aes(mixing_coefficient, mean_intra_clone_distances_clone1)) +
  geom_point() +
  theme_bw() +
  geom_smooth(method = "lm")

# ggsave("results/top50_cpg.pdf")
```

```{r}
df_homogeneity_similarity %>%
  filter(mixing_coefficient > 0.9, mean_intra_clone_distances_clone1 - mean_intra_clone_distances_clone2 > 5)

df_homogeneity_similarity %>%
  filter(mixing_coefficient > 0.9 & mixing_coefficient < 0.92, mean_intra_clone_distances_clone1 - mean_intra_clone_distances_clone2 < 2)

```



```{r}
df_homogeneity_similarity %>%
  arrange(desc(mean_intra_clone_distances))
```


```{r}
mean(as.matrix(df["B_131_5_22_52_2_3",]))
mean(as.matrix(df["B_278_8_8_261_1_1",]))
```


```{r}
c <- "B_278_8_8_261_1_1"
c <- "B_184_5_7_167_2_1"
c <- "B_2_5_22_312_1_3"
b <- "B_131_5_22_52_2_3"
b <- "B_250_10_1_375_2_3"
b <- "B_26_3_3_21_2_1"

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = list(
    "B_2_5_22_312_1_3" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == c, ]),
    "B_26_3_3_21_2_1" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == b, ])
  ),
  cols.highlight = c("blue", "red")
) + coord_fixed(ratio = 1)
```

```{r}
c <- "B_143_6_6_278_2_1"
b <- "B_2_5_22_312_1_3"

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = list(
    "B_143_6_6_278_2_1" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == c, ]),
    "B_2_5_22_312_1_3" = rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == b, ])
  ),
  cols.highlight = c("blue", "red")
) + coord_fixed(ratio = 1)
```


```{r}
distant_clones <- df_homogeneity_similarity %>%
  arrange(mixing_coefficient) %>%
  head(1) %>%
  tail(1) %>%
  select(clone1, clone2) %>%
  unlist()


distant_clones_highlight <- lapply(distant_clones, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(distant_clones_highlight) <- distant_clones

DimPlot(
  bcr_cpg,
  reduction = "UMAP",
  cells.highlight = distant_clones_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```

```{r}
close_clones <- df_homogeneity_similarity %>%
  arrange(desc(mixing_coefficient)) %>%
  head(1) %>%
  tail(1) %>%
  select(clone1, clone2) %>%
  unlist()


close_clones_highlight <- lapply(close_clones, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(close_clones_highlight) <- close_clones

DimPlot(
  bcr_cpg,
  reduction = "UMAP",
  cells.highlight = close_clones_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```



```{r}
df_homogeneity_similarity
hetereogeneous_clones <- df_homogeneity_similarity %>%
  arrange(desc(mean_intra_clone_distances_clone1)) %>%
  select(clone1) %>%
  unique() %>%
  head(5) %>%
  pull(clone1)

hetereogeneous_clones_highlight <- lapply(hetereogeneous_clones, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(hetereogeneous_clones_highlight) <- hetereogeneous_clones

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = hetereogeneous_clones_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```

```{r}
homogeneous_clones <- df_homogeneity_similarity %>%
  arrange(desc(mean_intra_clone_distances_clone1)) %>%
  select(clone1) %>%
  unique() %>%
  tail(5) %>%
  pull(clone1)

homogeneous_clones_highlight <- lapply(homogeneous_clones, function(x) rownames(bcr_cpg@meta.data[bcr_cpg@meta.data$clone_id == x, ]))
names(homogeneous_clones_highlight) <- homogeneous_clones

DimPlot(
  bcr_cpg,
  reduction = "umap",
  cells.highlight = homogeneous_clones_highlight,
  cols.highlight = c("blue", "red", "green", "yellow", "orange")
) + coord_fixed(ratio = 1)
```



Why do I get 1.02 for mixing score? 

What is different between clones that overlap very well? 

Correlation mixing coefficient and clone size and homogeneity


homogeneity between time points
Average distance between cells within a clone


For each clone, calculate the background by splitting the clone into 2?
Background: for each clone, compare similarity score with a random selection of cells


What happens if I go not symmetric? Only cells from c, not b?


Distribution of cell types across clones

```{r fig.width=10, fig.height=5}
bcr_cpg@meta.data %>%
  filter(clone_id %in% clones) %>%
  merge(cluster_anno, by.x = "clone_id", by.y = "row.names") %>%
  group_by(clone_id, y) %>%
  count(celltype) %>%
  ggplot(aes(fill=celltype, y=n, x=clone_id)) + 
    geom_bar(position="fill", stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  facet_grid(~y, scales = "free")
```

Get mean PC space for top50 clones
kmeans clustering 

```{r}
cells_df <- object@meta.data %>% rownames_to_column(var = "cell_id") %>% filter(clone_id %in% clones) %>% select(clone_id, cell_id)

embeddings <- Embeddings(object = object[[reduction]])[, dims]

clone_pc_means <- as.data.frame(embeddings) %>%
  rownames_to_column(var = "cell_id") %>%
  merge(cells_df, by="cell_id") %>%
  group_by(clone_id) %>%
  summarise(across(where(is.numeric), mean)) %>%
  as.data.frame()

rownames(clone_pc_means) <- clone_pc_means$clone_id
clone_pc_means$clone_id <- NULL

clone_pc_means_kmeans <- t(clone_pc_means)
colnames(clone_pc_means_kmeans) <- rownames(clone_pc_means)

set.seed(21)
clone_pc_means_kmeans <- kmeans(x = clone_pc_means, centers = 4)

```

```{r}
clone_pc_means_dist <- as.data.frame(as.matrix(dist(clone_pc_means)))
colnames(clone_pc_means_dist) <- clone_pc_means$clone_id
rownames(clone_pc_means_dist) <- clone_pc_means$clone_id
```

```{r}
clone_pc_means_dist_df <- clone_pc_means_dist %>%
  rownames_to_column(var = "clone1") %>%
  pivot_longer(cols = -clone1, names_to = "clone2", values_to = "pca_distance") %>%
  filter(pca_distance != 1)

df_homogeneity_similarity <- merge(df_homogeneity_similarity, clone_pc_means_dist_df, by=c("clone1", "clone2"))
```

```{r fig.height=4, fig.width=5}
df_homogeneity_similarity %>%
  ggplot(aes(mixing_coefficient, pca_distance, colour=abs(mean_intra_clone_distances_clone1 - mean_intra_clone_distances_clone2))) +
  theme(legend.title = element_blank())+#"abs(difference median distance")) +
  geom_point()
```


```{r}
bcr_cpg@meta.data %>%
  filter(clone_id %in% clones) %>%
  merge(clone_pc_means_kmeans$cluster, by.x = "clone_id", by.y = "row.names") %>%
  group_by(clone_id, y) %>%
  count(celltype) %>%
  ggplot(aes(fill=celltype, y=n, x=clone_id)) + 
    geom_bar(position="fill", stat="identity") +
    theme(axis.text.x = element_text(angle = 45, hjust=1)) +
  facet_grid(~y, scales = "free")
```


```{r}
table(merge(
  as.data.frame(clone_pc_means_kmeans$cluster),
  as.data.frame(cluster_anno),
  by = "row.names"
)[c("clone_pc_means_kmeans$cluster", "cluster_anno")]) %>%
  heatmap(scale = "none")
```




